---
title: "HR Analysis"
author: "Paul Chong"
date: "10/27/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r,include=FALSE}
library(tidyverse)
library(kableExtra)
library(xgboost)
library(SHAPforxgboost)
library(devtools)
library(caret)
library(plotly)
library(naniar)
library(forcatla)
library(scales)
library(lsr)
library(wesanderson)
library(e1071)
library(ggthemes)
```


```{r}
test <- read.csv("~/Documents/STAT472/aug_test.csv")
train <- read.csv("~/Documents/STAT472/aug_train.csv",header = TRUE, na.strings = c("","Unspecified"))
```

## Cleaning Data
```{r}
train <- train %>% 
  rename(relevant_experience = relevent_experience) %>% 
  mutate(relevant_experience = factor(ifelse(relevant_experience == "Has relevent experience", 'Yes', 'No'))) %>% 
  mutate(company_size = replace(company_size, company_size == '10/49', '10-49'))
```


## Transforming variables into factors

```{r pressure, echo=FALSE}
train <- train %>%
  mutate_if(is.character, as.factor) %>%
  mutate(city_development_index = factor(city_development_index)) %>%
  mutate(target = factor(target))
```



```{r}
train %>% 
  group_by(education_level, relevant_experience) %>% 
  summarise(count = n()) %>% 
  mutate(ratio = count / sum(count),
         label = percent(ratio %>% round(2))) %>%
  mutate(education_level = factor(education_level, 
                                  levels = c("Primary School",
                                             "High School",
                                             "Graduate",
                                             "Masters",
                                             "Phd"))) %>% 
  filter(education_level != 'NA') %>% 
  ggplot(aes(relevant_experience, count, fill=relevant_experience)) +
  geom_bar(stat='identity', color = 'black') +
  facet_wrap(~ education_level, scales = "free")+
  geom_label(aes(label=label), fill="#FFF9F5", colour = "black",
             fontface = "italic", vjust = 0.5) +
  theme_hc() +
  theme(plot.title=element_text(size=30, face="bold", hjust = 0.5),
        plot.subtitle=element_text(size=22, hjust = 0.5),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(vjust = 5, 
                                   hjust = 0.5, 
                                   size = 14),
        axis.text.y = element_text(size = 16),
        legend.position = "top",
        legend.title=element_blank(),
        legend.text=element_text(size=18),
        strip.background = element_rect(fill = 'white')) +
  scale_fill_manual(values = c('tomato', 'skyblue')) +
  theme(legend.position="none")+
  theme(axis.text.x = element_text(vjust = 1, 
                                   hjust = 0.5, 
                                   size = 12)) +
  theme(strip.text = element_text(colour = 'black', size = 18)) +
  labs(title = "Relationship between relevant experience and education level", size = 5)
```

We can see here that there is a positive relationship between the variables. The higher the education level, the greater the amount of candidates with relevant experience in the data science field.



## Distribution by Years

```{r}
train %>%
  group_by(experience) %>%
  summarise(count = n())%>%
  mutate(ratio = count / sum(count),
         label = percent(ratio %>% round(2))) %>% 
  mutate(experience = factor(experience, levels = c("<1", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", ">20"))) %>% 
  drop_na() %>% 
  ggplot(aes(x = experience, y = count)) +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  geom_bar(stat = "identity", position="identity", fill = "skyblue", color = 'black')+
  geom_label(aes(label=label), fill="#FFF9F5", colour = "black",
             fontface = "italic", vjust = 0.5, hjust = 0.3) +
  coord_flip()+
  theme(legend.position="none")+
  theme_hc() +
  theme(plot.title=element_text(size=30, face="bold", hjust = 0.5),
        plot.subtitle=element_text(size=22, hjust = 0.5),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 20),
        axis.ticks.y = element_blank(),
        legend.position = "top",
        legend.title=element_blank(),
        legend.text=element_text(size=18),
        strip.background = element_rect(fill = 'white')) +
  labs(title = "Experience distribution by years") +
  theme(panel.grid.major.x = element_line(color="#cbcbcb"), 
        panel.grid.major.y=element_blank())
```

## Experience vs Job Search
```{r}
train %>% 
  group_by(experience, target) %>%
  summarise(count = n())%>%
  mutate(ratio = count / sum(count)) %>% 
  mutate(experience = factor(experience, levels = c("<1", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", ">20"))) %>% 
  drop_na() %>% 
  ggplot(aes(experience, ratio, fill=target)) + 
  geom_bar(stat='identity', color = 'black') +
  theme_hc() +
  theme(plot.title=element_text(size=30, face="bold", hjust = 0.5),
        plot.subtitle=element_text(size=22, hjust = 0.5),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(vjust = 1, 
                                   hjust = 0.5, 
                                   size = 14),
        axis.text.y = element_text(size = 20),
        legend.position = "top",
        legend.title=element_blank(),
        legend.text=element_text(size=18),
        strip.background = element_rect(fill = 'white')) +
  scale_fill_manual(values = c('skyblue', 'tomato'),
                    labels = c("Not searching for a new job", 
                                 "Searching for a new job")) +
  coord_flip() +
  scale_y_continuous(labels = percent) +
  labs(title = "Relationship between experience and job search") +
  theme(panel.grid.major.x = element_line(color="#cbcbcb"), 
        panel.grid.major.y=element_blank())
```



## Cleaning for XGBoost preparation

```{r}
# Data transformation
train <- train %>% 
  select(-enrollee_id, -city) %>% 
  mutate(city_development_index = 
           as.numeric(as.character(city_development_index))) %>% 
  mutate(training_hours = 
           as.numeric(as.character(training_hours))) %>%
  mutate(target = 
           as.numeric(as.character(target)))
```

```{r}
# Label Encoding for qualitative variable
train$gender = 
  ifelse(train$gender == "Male", 1,
  ifelse(train$gender == "Female", 2, 0))

train$relevant_experience = 
  ifelse(train$relevant_experience == "Has relevant experience", 1, 0)

train$enrolled_university = 
  ifelse(train$enrolled_university == "no_enrollment",0,
  ifelse(train$enrolled_university == "Part time course", 2, 3))

train$education_level = 
  ifelse(train$education_level == "Primary School", 0,
  ifelse(train$education_level == "High School", 1,
  ifelse(train$education_level == "Graduate", 2,
  ifelse(train$education_level == "Masters", 3, 4))))

train$major_discipline = 
  ifelse(train$major_discipline == "No Major", 0,
  ifelse(train$major_discipline == "Other", 1,
  ifelse(train$major_discipline == "Arts", 2,
  ifelse(train$major_discipline == "Business Degree", 3,
  ifelse(train$major_discipline == "Humanities", 4, 5
         )))))

train$experience = 
  ifelse(train$experience == "<1",0,
  ifelse(train$experience == "1",1,
  ifelse(train$experience == "2",2,
  ifelse(train$experience == "3",3,
  ifelse(train$experience == "4",4,
  ifelse(train$experience == "5",5,
  ifelse(train$experience == "6",6,
  ifelse(train$experience == "7",7,
  ifelse(train$experience == "8",8,
  ifelse(train$experience == "9",9,
  ifelse(train$experience == "10",10,
  ifelse(train$experience == "11",11,
  ifelse(train$experience == "12",12,
  ifelse(train$experience == "13",13,
  ifelse(train$experience == "14",14,
  ifelse(train$experience == "15",15,
  ifelse(train$experience == "16",16,
  ifelse(train$experience == "17",17,
  ifelse(train$experience == "18",18,
  ifelse(train$experience == "19",19,
  ifelse(train$experience == "20",20,21
  )))))))))))))))))))))

train$company_size = 
  ifelse(train$company_size == "<10",0,
  ifelse(train$company_size == "10/49",1,
  ifelse(train$company_size == "50-99",2,
  ifelse(train$company_size == "10-500",3,
  ifelse(train$company_size == "500-999",4,
  ifelse(train$company_size == "1000-4999",5,
  ifelse(train$company_size == "5000-9999",6,7
  )))))))

train$company_type = 
  ifelse(train$company_type == "Early Stage Startup",0,
  ifelse(train$company_type == "Funded Startup",1,
  ifelse(train$company_type == "NGO",2,
  ifelse(train$company_type == "Public Sector",3,
  ifelse(train$company_type == "Pvt Ltd",4,5
  )))))

train$last_new_job = 
  ifelse(train$last_new_job == "never",0,
  ifelse(train$last_new_job == "1",1,
  ifelse(train$last_new_job == "2",2,
  ifelse(train$last_new_job == "3",3,
  ifelse(train$last_new_job == "4",4,5
  )))))


```



```{r}
# Split dependents from independent variables
y = train$target
x = data.matrix(train[,-12])

# Split data in train-validation
numberTrain = round(length(y) * .8)

train_data = x[1:numberTrain,]
train_labels = y[1:numberTrain]

val_data = x[-(1:numberTrain),]
val_labels = y[-(1:numberTrain)]

# Create Dmatrixs
dtrain = xgb.DMatrix(data = train_data, label= train_labels)
dval = xgb.DMatrix(data = val_data, label= val_labels)

dtrain
```






## Checking balance of "Target" variable for XGBoost
```{r}
train %>% 
  select(target) %>%
  mutate_if(is.numeric, as.factor) %>%
  group_by(target) %>%
  summarise(Count = n())%>%
  mutate(Ratio = Count / sum(Count),
         label = percent(Ratio %>% round(4))) %>%
  ggplot(aes(x = target, y = Count, fill=target)) +
  geom_bar(stat = "identity", color = "black")+
  geom_label(aes(label=label), fill="#FFF9F5", colour = "black",
             fontface = "italic", vjust = 0.5)+
  scale_fill_manual(values = c('skyblue', 'tomato')) +
  theme_hc() +
  theme(plot.title=element_text(size=30, face="bold", hjust = 0.5),
        plot.subtitle=element_text(size=22, hjust = 0.5),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(vjust = 5, 
                                   hjust = 0.5, 
                                   size = 14),
        axis.text.y = element_text(size = 16),
        legend.position = "top",
        legend.title=element_blank(),
        legend.text=element_text(size=18),
        strip.background = element_rect(fill = 'white')) +
  scale_x_discrete(labels=c("Not searching for a job", "Searching for a job")) +
  theme(legend.position="none")+
  labs(title = "Target distribution")
```

We can see that the distribution of the target is very imbalanced. 

```{r}
positives = sum(train_labels == 1)
negatives = sum(train_labels == 0)

params = list(max.depth = 6,
                  objective = "binary:logitraw",
                  eval_metric = "auc",
                  scale_pos_weight = negatives/positives,
                  eta = 0.1,
                  subsample = 0.7,
                  nthread = 16)
# create gridsearch for "eta" and "subsample" 
# Google search for hyper parameter for optimization
params
```



```{r}
set.seed(472)
xgbcv <- xgb.cv( params = params,
                 data = dtrain,
                 nrounds = 1000,
                 early_stopping_rounds = 100,
                 nfold = 5,
                 showsd = T,
                 maximize=T)

```


```{r}
model = xgboost(data = dtrain,
                 params = params,
                 nround = 30)
```


```{r}
pred = predict(model,dval)
err = mean(as.numeric(pred > 0.5) != val_labels)
print(paste("val-error=", err))
```


```{r}
prediction = ifelse (pred > 0.5,1,0)

cm = confusionMatrix(as.factor(prediction), as.factor(val_labels))

draw_confusion_matrix <- function(cm) {
  
  layout(matrix(c(1,1,2)))
  par(mar=c(2,2,2,2))
  plot(c(100, 345), c(300, 450), type = "n", xlab="", ylab="", xaxt='n', yaxt='n')
  title('CONFUSION MATRIX', cex.main=2)
  
  rect(150, 430, 240, 370, col='skyblue')
  text(195, 435, 'Not looking for a job', cex=1.2)
  rect(250, 430, 340, 370, col='tomato')
  text(295, 435, 'Looking for a job', cex=1.2)
  text(125, 370, 'Prediction', cex=1.3, srt=90, font=2)
  text(245, 450, 'Reality', cex=1.3, font=2)
  rect(150, 305, 240, 365, col='tomato')
  rect(250, 305, 340, 365, col='skyblue')
  text(140, 400, 'Not looking for a job', cex=1.2, srt=90)
  text(140, 335, 'Looking for a job', cex=1.2, srt=90)
  
  res <- as.numeric(cm$table)
  text(195, 400, res[1], cex=1.6, font=2, col='white')
  text(195, 335, res[2], cex=1.6, font=2, col='white')
  text(295, 400, res[3], cex=1.6, font=2, col='white')
  text(295, 335, res[4], cex=1.6, font=2, col='white')
  
  plot(c(100, 0), c(100, 0), type = "n", xlab="", ylab="", main = "DETAILS", xaxt='n', yaxt='n')
  text(10, 85, names(cm$byClass[1]), cex=1.2, font=2)
  text(10, 70, round(as.numeric(cm$byClass[1]), 3), cex=1.2)
  text(30, 85, names(cm$byClass[2]), cex=1.2, font=2)
  text(30, 70, round(as.numeric(cm$byClass[2]), 3), cex=1.2)
  text(50, 85, names(cm$byClass[5]), cex=1.2, font=2)
  text(50, 70, round(as.numeric(cm$byClass[5]), 3), cex=1.2)
  text(70, 85, names(cm$byClass[6]), cex=1.2, font=2)
  text(70, 70, round(as.numeric(cm$byClass[6]), 3), cex=1.2)
  text(90, 85, names(cm$byClass[7]), cex=1.2, font=2)
  text(90, 70, round(as.numeric(cm$byClass[7]), 3), cex=1.2)
  
  text(30, 35, names(cm$overall[1]), cex=1.5, font=2)
  text(30, 20, round(as.numeric(cm$overall[1]), 3), cex=1.4)
  text(70, 35, names(cm$overall[2]), cex=1.5, font=2)
  text(70, 20, round(as.numeric(cm$overall[2]), 3), cex=1.4)
}  

draw_confusion_matrix(cm)
```

```{r}
xgb_importance = xgb.importance(names(x), model = model)

xgb_importance %>% head(10) %>% kbl('html', 
    caption = 'Importance of variables') %>%
  kable_styling(bootstrap_options = c("striped", 
                                      "hover"))
```








