---
title: "472 Project"
author: "Ali Almarbooei"
date: "11/05/2021"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the Lib and data
```{r} 
library(ggplot2)
library(gridExtra)
library(dplyr)
library(e1071)

```


```{r}
#load the data 
df = read.csv("aug_train.csv", na.strings=c("","Unspecified"))
head(df)
```



## Explore data
```{r}
#Check gender arrangement 
barplot(table(df$gender),main="Gender",xlab="Gender",ylab="Count",border="red",col="blue",density=10)

```
```{r}

barplot(table(df$relevent_experience),main="Experience",xlab="Experience Level",ylab="Count",border="red",col="blue",density=10)

```
```{r}

barplot(table(df$enrolled_university),main="University Enrollment",xlab="University Enrollment",ylab="Count",border="red",col="blue",density=5)

```
```{r}

barplot(table(df$education_level),main="Education",xlab="Education",ylab="Count",border="red",col="blue",density=5)

```
```{r}

barplot(table(df$major_discipline),main="Major Discipline",xlab="Discipline",ylab="Count",border="red",col="blue",density=5)

```

```{r}
barplot(table(df$last_new_job),main="Major Discipline",xlab="Discipline",ylab="Count",border="black",col="red",density=5)
```
```{r}
barplot(table(df$company_size),main="Company Size",xlab="size",ylab="Count",border="black",col="red",density=5)
```
```{r}
barplot(table(df$company_type),main="Company Type",xlab="Company type",ylab="Count",border="black",col="red",density=5)
```

## Unspecified and Factors
```{r}
## Replace to missing value with unspecified
df = df %>%
  mutate(gender = replace(gender,
                          is.na(gender),
                          "Unspecified"),
         experience = replace(experience,
                              is.na(experience),
                              "Unspecified"),
         major_discipline = replace(major_discipline,
                                    is.na(major_discipline),
                                    "Unspecified"),
         education_level = replace(education_level,
                                   is.na(education_level),
                                   "Unspecified"),
         enrolled_university = replace(enrolled_university,
                                       is.na(enrolled_university),
                                       "Unspecified"),
         company_size = replace(company_size,
                                       is.na(company_size),
                                       "Unspecified"),
         company_type = replace(company_type,
                                       is.na(company_type),
                                       "Unspecified"),
         last_new_job = replace(last_new_job, 
                                is.na(last_new_job),
                                "Unspecified"))
head(df)
```

```{r}
#changing data from character to factors
df = subset(df, select = -c(enrollee_id))
# Change Factor
df$gender = as.factor(df$gender)
df$last_new_job = as.factor(df$last_new_job)
df$relevent_experience = as.factor(df$relevent_experience)
df$experience =as.factor(df$experience)
df$major_discipline = as.factor(df$major_discipline)
df$enrolled_university = as.factor(df$enrolled_university)
#df$company_size = as.factor(df$company_size)
df$company_type = as.factor(df$company_type)
df$target = as.factor(df$target)

summary(df)
```



## encode data 
```{r}
#function to encode the data
encodeData <- function(x, order = unique(x)) {
  x <- as.numeric(factor(x, levels = order, exclude = NULL))
  x
}
```

```{r}

```

## view the distribution based on new numerical categories
```{r}
# print("Gender: ")
summary(as.factor(df$gender))
print("Last new Job")
summary(as.factor(df$last_new_job))
print("Education Level")
summary(as.factor(df$education_level))
print("Experience: ")
summary(as.factor(df$experience))
print("Relevant Experience")
summary(as.factor(df$relevent_experience))
print("Major Discipline: ")
summary(as.factor(df$major_discipline))
print("Enrolled University?")
summary(as.factor(df$enrolled_university))
print("Company size: ")
summary(as.factor(df$company_size))
print("Company type")
summary(as.factor(df$company_type))
```
in most of the variables the ranking has been done based on the size of each category. the one with large category comes first

```{r}
#factorize categorical data in specific order
df$education_level = dplyr::case_when(df$education_level=="Primary School"~1,
                                              df$education_level=="High School"~2,
                                              df$education_level=="Graduate"~3,
                                              df$education_level=="Masters"~4,
                                              df$education_level=="Phd"~5,
                                              df$education_level=="Unspecified"~6,
                                              )

df$company_size = dplyr::case_when(df$company_size=="Unspecified"~0,
                                  df$company_size=="<10"~1,
                                  df$company_size=="10/49"~2,
                                  df$company_size=="50-99"~3,
                                  df$company_size=="100-500"~4,
                                  df$company_size=="500-999"~5,
                                  df$company_size=="1000-4999"~6,
                                  df$company_size=="5000-9999"~7,
                                  df$company_size=="10000+"~8,)
```

```{r}
#since the city has many categorical values it cannot be factored. so we remove it 
df$city=NULL

```





#Split data into training and testing datasets
```{r}
n = nrow(df)
train_set <- 1:(n * 0.6)
train <- df[train_set,]     #60% of data for training 
test <- df[-train_set,]     #40% of data for testing
```

## Create a logistical regression model
```{r}

## GLM 
set.seed(456)
glm.fit = glm(formula = target ~ ., data = train,family = "binomial")
summary(glm.fit)
```
Based on the above logistic model results there are seven variables that are statistically significant: city development index, gender, relevant_experience, enrolled_university, education_level, company size and company type.
major_discipline, experience, last_new_job and training hours are not significant for study hence we except them from further study

```{r}
#calculating confidence interval
#confint(glm.fit)
```


## Creating model using significant variables ------------------ 6 variables ---------------------------------
```{r}
my4model <- glm(target ~ city_development_index + experience + training_hours + gender + company_size + company_type, data = train, family = "binomial")
summary(my4model)
```

```{r}
#confidence interval
#confint(my4model)
```


### checking if the predictor variable are significant
```{r}
#Wald test
library(aod)
wald.test(b = coef(my4model), Sigma = vcov(my4model), Terms = 4:6)
```
## -----------------------------------------------------------------------------


##--------------------------- Histograms ---------------------------------------
```{r}
Target_num = as.numeric(train$target) - 1
library(popbio)
```

## Training Hours 
```{r}
logi.hist.plot(as.numeric(train$training_hours), Target_num, logi.mod = 1, type = "hist",
  boxp = FALSE, rug = FALSE, ylabel = "Job Change Probability",
  ylabel2 = "", xlabel = "Hours", mainlabel = "Training Hours", las.h = 1,  counts = TRUE, col = " light blue")
```
Most of the jobs have less than 150 training hours, it happen that these jobs also have larger number of those who want to change their job
There is slight decrease in probability to change job  as training hours increases. 

## City Developmenx index
```{r}
logi.hist.plot(train$city_development_index, Target_num, logi.mod = 1, type = "hist",
  boxp = FALSE, rug = FALSE, ylabel = "Job Change Probability",
  ylabel2 = " ", xlabel = " index", mainlabel = " City Development Index", las.h = 1,
  counts = TRUE, col = "light blue")
```
Most people lives in cities with higher development as it can be seen from the proportion of people in higher city index 
However, the probability of people wanting to change descreases with increase in city development index

## experience
```{r}
experience_num1 = as.numeric(train$experience) 

logi.hist.plot(experience_num1, Target_num, logi.mod = 1, type = "hist",
  boxp = FALSE, rug = FALSE, ylabel = "Job Change Probability",
  ylabel2 = " ", xlabel = " Years", mainlabel = " experience", las.h = 1,
  counts = TRUE, col = "blue") 
```
the is high number of people without experience, there is also high number of people who have experience of between 10 and 22 years
the probability to change job increases slightly with increase in years of experience

## education level 
```{r}
education_level1 = train$education_level
logi.hist.plot(education_level1, Target_num, logi.mod = 1, type = "hist",
  boxp = FALSE, rug = FALSE, ylabel = "Job Change Probability",
  ylabel2 = " ", xlabel = "1:Primary School 2:Highschool 3:Graduate   4: Masters     5: PHD   6:Unspecified", mainlabel = " Education Level", las.h = 1,
  counts = TRUE, col = "light blue")
```
Education level does not seem to be a big factor of people wanting to change their current job


## company size 
```{r}

logi.hist.plot(train$company_size, Target_num, logi.mod = 1, type = "hist",
  boxp = FALSE, rug = FALSE, ylabel = "Job Change Probability",
  ylabel2 = " ", xlabel = " size", mainlabel = " Company size", las.h = 1,
  counts = TRUE, col = "light blue") 

```
people is small companies tends to want to want to move to other company
##------------------------------------------------------------------------------
##?logi.hist.plot 



## Accuracy 
```{r}
# Check for accuracy
prob = predict(glm.fit, test,type = "response") 
glm.pred <- rep(0, length(prob))
glm.pred[prob > 0.5] <- 1

mean(glm.pred == test$target)

```
```{r}
library(caret)
cm <- confusionMatrix(as.factor(glm.pred),test$target)
```

```{r}
draw_confusion_matrix <- function(cm) {

  layout(matrix(c(1,1,2)))
  par(mar=c(2,2,2,2))
  plot(c(100, 345), c(300, 450), type = "n", xlab="", ylab="", xaxt='n', yaxt='n')
  title('CONFUSION MATRIX', cex.main=2)

  # create the matrix 
  rect(150, 430, 240, 370, col='blue')
  text(195, 435, 'Positive (Change job)', cex=1.2)
  rect(250, 430, 340, 370, col='purple')
  text(295, 435, 'Negative (Keep current Job)', cex=1.2)
  text(125, 370, 'Predicted', cex=1.3, srt=90, font=2)
  text(245, 450, 'Actual', cex=1.3, font=2)
  rect(150, 305, 240, 365, col='purple')
  rect(250, 305, 340, 365, col='blue')
  text(140, 400, 'Positive', cex=1.2, srt=90)
  text(140, 335, 'Negative', cex=1.2, srt=90)

  # add in the cm results 
  res <- cm$table
  text(195, 400, res[1], cex=1.6, font=2, col='white')
  text(195, 335, res[2], cex=1.6, font=2, col='white')
  text(295, 400, res[3], cex=1.6, font=2, col='white')
  text(295, 335, res[4], cex=1.6, font=2, col='white')
  
  # add in the specifics 
  plot(c(100, 0), c(100, 0), type = "n", xlab="", ylab="", main = "DETAILS", xaxt='n', yaxt='n')
  text(10, 85, names(cm$byClass[1]), cex=1.2, font=2)
  text(10, 70, round(as.numeric(cm$byClass[1]), 3), cex=1.2)
  text(30, 85, names(cm$byClass[2]), cex=1.2, font=2)
  text(30, 70, round(as.numeric(cm$byClass[2]), 3), cex=1.2)
  text(50, 85, names(cm$byClass[5]), cex=1.2, font=2)
  text(50, 70, round(as.numeric(cm$byClass[5]), 3), cex=1.2)
  text(70, 85, names(cm$byClass[6]), cex=1.2, font=2)
  text(70, 70, round(as.numeric(cm$byClass[6]), 3), cex=1.2)
  text(90, 85, names(cm$byClass[7]), cex=1.2, font=2)
  text(90, 70, round(as.numeric(cm$byClass[7]), 3), cex=1.2)

  # add in the accuracy information 
  text(30, 35, names(cm$overall[1]), cex=1.5, font=2)
  text(30, 20, round(as.numeric(cm$overall[1]), 3), cex=1.4)
  text(70, 35, names(cm$overall[2]), cex=1.5, font=2)
  text(70, 20, round(as.numeric(cm$overall[2]), 3), cex=1.4)
  
}  
```

```{r}
draw_confusion_matrix(cm)
```
# code draw confusion matrix function adopted from https://newbedev.com/r-how-to-visualize-confusion-matrix-using-the-caret-package
https://stats.idre.ucla.edu/r/dae/logit-regression/